Step-by-Step: Training IMU-based Gestures with Live Feedback

### 2023-12-12 docs/latex/paper.tex
final paper commit

@@

@@

The window slides forward one frame at a time. If the most recent frame in a window is labelled with a gesture, the window is labelled with that gesture. Otherwise, the window is labelled with no gesture. As new windows are collected, the model is trained on the windows using gradient descent. Training is iterative and the model's training data is updated in realtime, allowing the user to receive live feedback on the model's performance.

@@

Along with the groundtruth labels, the model's predictions are used to generate live feedback for the user using both auditory and visual cues. The auditory cue is a simple synthesizer that plays a tone corresponding to the model's prediction. The visual cue is a simple GUI that displays the IMU timeseries, model's predictions, and groundtruth labels in realtime.

@@

\section{Results}
After just a few examples of a gesture, the model converges on solution that reliably distinguishes gesture from non-gesture. Performance improves as more examples are provided. Here, the authors only test the model's ability to recognize a single gesture, but based on prior experience using this model in non-realtime training scenarios, the authors expect the model to be able to distinguish multiple gestures with high accuracy.
% The model's performance is shown in Figure \ref{fig:performance}.

@@

\bibliography{example_paper}
\bibliographystyle{mlsys2024}

>>>>

\begin{figure}[h]
    \includegraphics[width=0.5\textwidth]{../figures/model_diagram.jpg}
    \caption{Neural Network Architecture. The sensor branch (left) receives 20 frames (200 ms at 100 Hz) of 3-channel (x, y, z) IMU data. The sensor branch feeds into a 1D-convolutional layer with 8 filters, each also 20 frames in width. The memory branch (right) receives 15 frames of 1-hot encoded prediction data. The two branches are concatenated and fed into a fully connected output layer to predict the currently active gesture.}
    \label{fig:your_label}
\end{figure}

@@

\begin{figure*}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{../figures/realtime_training.jpg}
    \caption{Live Learning Feedback: The user performs a gesture and receives live feedback on the model's performance. Bottom Panel: 10 seconds of training data. Acceleration components (x, y, z) are shown in color; magnitude is shown in grey. Each magnitude peak is marked by a thin vertical line, indicating it has been labelled as a gesture in the training data. Top Panel: 10 seconds of inference data. At t = 0 seconds the neural network is completely untrained. Naive to what is and is not a gesture, it predicts gesture and non-gesture with equal probability (0.5). As the model trains on non-gesture date, it begins to predict non-gesture with higher probability. At t = 5 seconds, the model has converged on a solution that reliably distinguishes non-gesture. At t = 10 seconds, the model has converged on a solution that reliably distinguishes gestures. Given more training data and time to learn, the model's performance would continue to improve with gessture and non-gesture probabilities approaching 1 and 0, respectively. See supplementary video for live demo in the github repo.}
    \label{fig:realtime_training}
\end{figure*}

@@

The window slides forward one frame at a time. If the most recent frame in a window is labelled with a gesture, the window is labelled with that gesture. Otherwise, the window is labelled with no gesture. As new windows are collected, the model is trained on the windows using gradient descent. Training is iterative and the model's training data is updated in realtime, allowing the user to receive live feedback on the model's performance. This process is illustated in Figure \ref{fig:realtime_training}.

@@

Along with the groundtruth labels, the model's predictions are used to generate live feedback for the user using both auditory and visual cues. The auditory cue is a simple synthesizer that plays a tone corresponding to the model's prediction. The visual cue is a simple GUI that displays the IMU timeseries, model's predictions, and groundtruth labels in realtime. The visual elements are illustated in Figure \ref{fig:realtime_training}.

@@

\section{Results}
After just a few examples of a gesture, the model converges on solution that reliably distinguishes gesture from non-gesture. Performance improves as more examples are provided. Here, the authors only test the model's ability to recognize a single gesture, but based on prior experience using this model in non-realtime training scenarios, the authors expect the model to be able to distinguish multiple gestures with high accuracy. A representative example of the model learning to recognize a single gesture in just 10 seconds is shown in Figure \ref{fig:realtime_training}.
% The model's performance is shown in Figure \ref{fig:performance}.

@@

% \bibliography{example_paper}
% \bibliographystyle{mlsys2024}
